Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-10-23 text/supp.tex
Clarify why we left body size out of model in SOM
Clean up paragraph in SOM

We fitted a multilevel beta regression model to the predicted probability of
heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed
dynamics. The beta distribution is useful when response data range on
a continuous scale between zero and one. We used a logit link function as is
typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,\\
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$
represents the predicted value for population $i$, class $j$, order $k$, and
species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$
represents a vector of predictors (such as lifespan) for population $i$ with
associated $\beta$ coefficients. The intercepts are allowed to vary from the
overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$),
taxonomic order ($\alpha^\mathrm{order}_k$), and species
($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \;
  \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \;
  \mathrm{species}}$. Where possible, we also allowed for error distributions
around the predictors by incorporating the standard deviation of the posterior
samples for the Gompertz parameters $\lambda$, $b$, and $\sigma$ around the
mean point value as normal distributions (not shown in the above equation). We
log transformed $\sigma$, time-series length, and lifespan to match the way
they are visually represented in Fig.~\ref{fig:correlates} and to make the
relationship approximately linear on the logit-transformed response scale. All
input variables were standardized by subtracting their mean and dividing by two
standard deviations \citep{gelman2008c} to make their coefficients comparable
in magnitude.

>>>>

We fitted a multilevel beta regression model to the predicted probability of
heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed
dynamics. The beta distribution is useful when response data range on
a continuous scale between zero and one. We used a logit link function as is
typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,\\
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$
represents the predicted value for population $i$, class $j$, order $k$, and
species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$
represents a vector of predictors (such as lifespan) for population $i$ with
associated $\beta$ coefficients. The intercepts are allowed to vary from the
overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$),
taxonomic order ($\alpha^\mathrm{order}_k$), and species
($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \;
  \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \;
  \mathrm{species}}$. Where possible, we also allowed for error distributions
around the predictors by incorporating the standard deviation of the posterior
samples for the Gompertz parameters $\lambda$, $b$, and $\sigma$ around the
mean point value as normal distributions (not shown in the above equation).

We log transformed $\sigma$, time-series length, and lifespan to match the way
they are visually represented in Fig.~\ref{fig:correlates} and to make the
relationship approximately linear on the logit-transformed response scale. All
input variables were standardized by subtracting their mean and dividing by two
standard deviations to make their coefficients comparable in magnitude
\citep{gelman2008c}. We excluded body length as a covariate because it was
highly correlated with lifespan, and lifespan exhibited more overlap across
taxonomy than body length. Lifespan is also more directly related to time and
potential mechanisms driving black swan dynamics.
